### 2014-03-21 figures/arch.tex
figure

\documentclass{article}
\usepackage[latin1]{inputenc}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,arrows,shadows,positioning,backgrounds,fit} %,external
%\tikzexternalize
\usepackage[active,pdftex,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\begin{document}
\sffamily
\pagestyle{empty}
\renewcommand{\rmdefault}{phv} % Arial
\renewcommand{\sfdefault}{phv} % Arial

\tikzstyle{component} = [diamond, draw,text badly centered,inner sep=0pt]
\tikzstyle{block} = [rectangle, draw,text centered, rounded corners,minimum size=1cm]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse]
\tikzstyle{nodeEdge} = [black,thick,->];
\tikzstyle{nodeBi} = [black,thick,<->];

\def\blockdist{2.3}

\begin{tikzpicture}[text width=6em,auto]

    % Place nodes
    \node [block] (core) {numerical engine};
    \node [block] (controller) [right=0.5cm of core] {model controller};
    \node [block] (schema) [below=0.5cm of core] {schematization};
    \node [block] (feed) [above=1cm of controller] {data feed};
    \node [block] (broker) [below right=0cm and 0.5cm of controller] {event+action\\ cache};
    \node [block] (cache) [above right=0cm and 0.5cm of controller] {data cache};
    \node [block] (maps) [right=0.5cm of cache] {time+map \\ server};
    \node [block] (sockets) [right=0.5cm of broker] {websocket server};
    \node [block] (browser) [below right=0 cm and 1cm of maps] {browser};

    % Connections
    \draw ($(schema.north)$) edge[nodeEdge] ($(core.south) $);
    \draw ($(core.east) $) edge[nodeBi] ($(controller.west)$);
    \draw ($(feed.south) $) edge[nodeEdge] ($(controller.north)$);
    \draw ($(feed.south) $) edge[nodeEdge] ($(controller.north)$);
    \draw ($(controller.north) $) edge[nodeEdge,out=90,in=180] ($(cache.west)$);
    \draw ($(controller.south) $) edge[nodeBi,out=270,in=180] ($(broker.west)$);
    \draw ($(cache.east) $) edge[nodeEdge] ($(maps.west)$);
    \draw ($(broker.east) $) edge[nodeBi] ($(sockets.west)$);
    \draw ($(maps.east) $) edge[nodeEdge,out=0,in=90] ($(browser.north)$);
    \draw ($(sockets.east) $) edge[nodeBi,out=0,in=270] ($(browser.south)$);

\end{tikzpicture}
\end{document}